---
title: "R Notebook"
output: html_notebook
---

First I made sure to library some commonly used packages in this course. Not only will this allow me to create graphs, but I will be able to code in the correct syntax.

```{r, message= FALSE}
library(DataComputing)
library(tidyverse)
library(mosaic)
library(ggplot2)
library(maps)
library(party)
```

I accessed my data from Kaggle. The [dataset](https://www.kaggle.com/karangadiya/fifa19) involves all of the FIFA players from FIFA 19.

### Upload Data Sets
```{r, message= FALSE}
###Using file.choose we store the csv address in dataset
Orig_Fifa_data<-read_csv("/Users/davidweiss/Desktop/Fall 2019/Stat 184/Final Project/FIFA DATA.csv")

head(Orig_Fifa_data)
```
Once the data was uploaded it was extremely messy.  There were random columns like ID and Photo that I wanted to get rid of. Also, variables like wage and height were character strings instead of numeric. 

```{r}
### Clean data so easier to work with

Cleaned_fifa_data<-
  Orig_Fifa_data %>%
  ### Need to make certain variables numeric
  mutate(Weight=parse_number(Weight),
         Value=parse_number(Value),
         Wage=parse_number(Wage),
         ### Regex to get height into inches
         Height=sapply(strsplit(sub("(\\d+)'(\\d+).*", "\\1-\\2", x = Height), "-"), function(x)
       as.numeric(x[1]) * 30.48 + as.numeric(x[2]) * 2.54)) %>%
  rename("Foot"="Preferred Foot") %>%
 ### Certain nationalities won't match when I join it with my second data set. 
  mutate(Nationality=if_else(Nationality == "China PR","China", Nationality)) %>%
  mutate(Nationality=if_else(Nationality %in% c("England","Wales", "Scotland", "Northern Ireland"),"UK", Nationality)) %>%
  mutate(region=if_else(Nationality=="Bosnia Herzegovina","Bosnia and Herzegovina", Nationality)) %>%
  mutate(Nationality=if_else(Nationality == "Republic of Ireland","Ireland", Nationality)) %>%
  mutate(Nationality=if_else(Nationality %in% c("United States", "United States of America"),"USA", Nationality)) %>%
  mutate(Nationality=if_else(Nationality=="Korea Republic","South Korea", Nationality))
  

### Select the relevant variables 
Cleaned_fifa_data<-
  Cleaned_fifa_data %>%
  select("Name", "Age", "Nationality", "Overall", "Potential", "Club", "Value", "Wage", "Foot", "Weak Foot", "Skill Moves", "Work Rate", "Position", "Height", "Weight", "Crossing",	"Finishing",	"HeadingAccuracy",	"ShortPassing",	"Volleys",	"Dribbling", "Curve",	"FKAccuracy",	"LongPassing",	"BallControl",	"Acceleration",	"SprintSpeed",	"Agility",	"Reactions",	"Balance",	"ShotPower",	"Jumping",	"Stamina",	"Strength",	"LongShots",	"Aggression",	"Interceptions",	"Positioning",	"Vision",	"Penalties",	"Composure",	"Marking",	"StandingTackle",	'SlidingTackle',	"GKDiving",	"GKHandling",	"GKKicking",	"GKPositioning",	"GKReflexes")

Cleaned_fifa_data
  
```
After looking through this data table I can see it looks a lot better and easier to work with.

The first thing I wanted to look at was where the top FIFA players were from. I thought it would be cool to build a function that would allow a user to select a number, representative of the top x players based on their overall, and then create a map to visualize where they are from. First I ranked the players based on their overall. Then I grouped by their region ie Nationality. This creates a data table that has the count of the players in one column and the Nationality in the other. Then I read in the map data and used a left join so all the cases in the Map are still there. Finally I used ggplot code to create the Map. I then ran it for the top 1000 players. 

```{r}
top_soccer_function <-function(x){
  
  Top<-
  Cleaned_fifa_data%>%
    ### make sure desc overall so the highest overall is the lowest rank
  mutate(rank=rank(desc(Overall))) %>%
  filter(rank<x) %>%
    ### Easier for join 
  mutate("region"=Nationality)
  
  Grouped_top <-
  Top %>%
  ### Group them by their nationality for the graph
  group_by(region) %>%
  summarise(count=n())
  
  ### Read in second data table
  World_map<- map_data("world")
  
  ### Join them
  Joined_top <-
  World_map %>%
  left_join(Grouped_top)
 
  ### Create the map
  Map<-
    ggplot(Joined_top, aes(x = long, y = lat, group=group)) +
    geom_polygon(aes(fill=count), colour = "white")+
    scale_fill_viridis_c(option = "C")
  
  ### Produce the Map when the function is run
  return(Map)
}
top_soccer_function(1000)
```


From here on out I wanted to work with the top 1000 players. Knowing each of the top 5 leagues has around 20 teams and 11 players start each game, 1000 gives us a pretty good breadth of talent across the board.

```{r}
Top<-
  Cleaned_fifa_data%>%
    ### make sure desc overall so the highest overall is the lowest rank
  mutate(rank=rank(desc(Overall))) %>%
  filter(rank<1000)
```


I also wanted to look at the different position groups and see if things like wage varied across the group. Furthermore, most clubs play a 4-3-3 with a goalie, so I wanted to see how the distribution of the player groups looked. 

```{r}
Position_data <-
  Top %>%
  ### Change each position into one of three categories
  mutate(Position=if_else(Position %in% c("ST", "LF", "RF", "LW", "RW", "RS", "LS", "CF"), "Attack", Position)) %>%
  mutate(Position=if_else(Position %in% c("CM", "LM", "RM", "LCM", "RCM", "CAM", "CDM", "LDM", "RDM", "LAM", "RAM"), "Midfield", Position)) %>%
  mutate(Position=if_else(Position %in% c("CB", "LB", "RB", "LWB", "RWB", "LCB", "RCB", "GK"), "Defense", Position))

Position_data
```


```{r}
### This will create a data table with each position and their count
Grouped_Position_data<-
  Position_data %>%
  group_by(Position) %>%
  summarise(count=n())
Grouped_Position_data

### Create a boxplot 
ggplot(data=Grouped_Position_data,aes(x=Position,y=count ))+
  geom_bar(stat='identity',position='stack', width=.9) 
```

I am actually quite surprised by this graphic. I figured FIFA may tailor their ratings to attacking players. I also thought the distribution may be a little more equal than this, even with Defense getting a boost from GK being considered a defender. However, maybe they consider players that play both attack and midfield midfielders. 


```{r}
ggplot( data = Position_data, aes(x = Age)) + 
  geom_density() + 
  aes(colour = Position) + 
  ### Add jitter to see actual distribution
  geom_point(alpha = 0.5, aes(y = 0.02),  
             position = position_jitter(height = 0.02)) 
```

This graph makes a lot of sense. We would expect the best players to be in their prime and that usually is correlated with the mid 20s. Interestingly we see some really old defenders and really young defenders in the jitter. Defenders can afford to be on the slower side, so them being old makes sense. However, the younger ones surprise me but seem like outliers as we see the distribution lines are much higher for attack and midfield at the younger ages.


```{r}
### Make a scatter plot showing wage and Overall
ggplot(data=Position_data,aes(x=Overall,y=Wage))+
  geom_point()+aes(colour=Position) 
```


Finally, I was interested to see if a players wage is correlated with their overall FIFA rating.  As seen in the plot above there is a clear upward trend. Also, we see that Attackers generally make more at a similar overall and defenders make the least. This makes sense as fans want to see goals and the top attackers make the largest impact. 
